---
title: "msia 400 lab 1 Finn Qiao"
output: html_notebook
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/finn/MSIA/msia400-finn")
```

```{r}
wine_df <- read.table('redwine.txt', header = TRUE)
dim(wine_df)
head(wine_df)
```

Problem 1
```{r}
mean(wine_df$RS, na.rm = TRUE)
mean(wine_df$SD, na.rm = TRUE)
```

Problem 2
```{r}
SD_omit <- wine_df$SD[!is.na(wine_df$SD)]
FS_omit <- wine_df$FS[!is.na(wine_df$SD)]
lmfit <- lm(SD_omit ~ FS_omit)
summary(lmfit)
```

Problem 3
```{r}
missing_SD <- subset(wine_df, complete.cases(wine_df$SD) == FALSE)
missing_SD$SD <- predict(lmfit, missing_SD)
missing_SD
```

```{r}
missing_SD_rows <- as.numeric(rownames(missing_SD))
wine_df_imputed <- wine_df
wine_df_imputed[missing_SD_rows,] <- missing_SD
mean(wine_df_imputed$SD)
```

Problem 4
```{r}
avg.imp <- function(a,avg) {
  missing <- is.na(a)
  imputed <- a
  imputed[missing] <- avg
  return (imputed)
}

wine_df_imputed$RS <- avg.imp(wine_df$RS, mean(wine_df$RS, na.rm = TRUE))
mean(wine_df_imputed$RS)
```

Prblem 5
```{r}
winemodel <- lm(QA ~ FA + VA + CA + RS + CH + FS + SD + DE + PH + SU + AL, wine_df_imputed)
summary(winemodel)
```

Problem 6

Based on the coefficients above, the greatest p value belongs to PH. It has a p value of 0.414480 and is highly statistically insignificant.

Problem 7
```{r}
library('DAAG')
```

```{r}
cv.lm(wine_df_imputed, winemodel, m=5)
```
Mean square is 0.426.

Problem 8
```{r}
ph_mean <- mean(wine_df_imputed$PH)
ph_sd <- sd(wine_df_imputed$PH)
```

```{r}
ph_lower_bound <- ph_mean - 3 * ph_sd
ph_upper_bound <- ph_mean + 3 * ph_sd
```

```{r}
redwine2 <- wine_df_imputed[with(wine_df_imputed, !(PH < ph_lower_bound | PH > ph_upper_bound)),]
dim(redwine2)
```

19 rows were dropped.

Problem 9
```{r}
winemodel <- lm(QA ~ FA + VA + CA + RS + CH + FS + SD + DE + PH + SU + AL, redwine2)
summary(winemodel)
```

The five most significant attributes seem to be VA, CH, SD, SU, and AL.

The previous model with the outliers included seemed to have more significant predictors. The R^2 and F statistic are slightly higher for this new model which suggests a slightly better fit.
