---
title: "msia 400 lab 1 Finn Qiao"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/finn/MSIA/msia400-finn")
```

```{r}
install.packages("expm")
library("expm")
```

# Problem 1

```{r}
trans_matrix <- read.table('markov100.txt', header=FALSE)
trans_matrix = as.matrix(trans_matrix)
```

```{r}
base_state = c(1, rep(0,99))
current_state = base_state
```

1a) Probability of being in state 5 after 10 transitions.
```{r}
prob10 = current_state %*% (trans_matrix %^% 10)
prob10[5]
```

1b) State 10 after 10 transitions.
```{r}
current_state = base_state
current_state[c(1,2,3)] = 1/3
prob10 = current_state %*% (trans_matrix %^% 10)
prob10[10]
```

1c) Steady state probability of being in state 1.
```{r}
A <- t(trans_matrix - diag(100))
A <- rbind(A, rep(1, 100))
b <- c(rep(0,100), 1)
steady <- qr.solve(A, b)
steady[1]
```

1d) Mean first passage time from state 1 to state 100.
```{r}
B = trans_matrix[1:99,1:99]
Q = diag(99) - B
e = c(rep(1,99))
m = solve(Q) %*% e
m[1]

```
# Problem 2

```{r}
web_matrix <- read.table('webtraffic.txt', header=TRUE)
web_matrix <- as.matrix(web_matrix)
```

2a) Traffic matrix that counts total traffic.
```{r}
Traffic <- colSums(web_matrix)
Traffic <- matrix(Traffic, nrow = 9, ncol = 9, byrow = TRUE)
Traffic
colSums(web_matrix)
```

2b) Construct probability matrix.
```{r}
Traffic[9,1] <- 1000
Traffic
```

```{r}
prob_matrix <- Traffic / rowSums(Traffic)
prob_matrix
```

2c) Steady state matrix.
```{r}
A <- t(prob_matrix - diag(9))
A <- rbind(A, rep(1, 9))
b <- c(rep(0,9), 1)
steady <- qr.solve(A, b)
steady
```

2d) Average time spent on website.

mean first passage time to the end from all other pages
```{r}
B = prob_matrix[1:8,1:8]
Q = diag(8) - B
e = c(0.1, 2, 3, 5, 5, 3, 3, 2)
m = solve(Q) %*% e
m[1]
```

2e) 
```{r}
prob_matrix
```

30% of state 2 to state 3 moves to state 2 to state 6
20% of state 2 to state 4 moves to state 2 to state 7

```{r}
new_prob_matrix <- prob_matrix
current23 <- new_prob_matrix[2,3]
current24 <- new_prob_matrix[2,4]

increase26 <- current23 * 0.3
increase27 <- current24 * 0.2

new_prob_matrix[2,3] <- new_prob_matrix[2,3] * 0.7
new_prob_matrix[2,6] <- new_prob_matrix[2,6] + increase26
new_prob_matrix[2,4] <- new_prob_matrix[2,4] * 0.8
new_prob_matrix[2,7] <- new_prob_matrix[2,7] + increase27

A <- t(new_prob_matrix - diag(9))
A <- rbind(A, rep(1, 9))
b <- c(rep(0,9), 1)
new_steady <- qr.solve(A, b)
new_steady
```

```{r}
steady
```

```{r}
var(steady)
var(new_steady)
```

The variance of the new steady state decreased and it seems like the traffic across the different pages are more balanced out after adding the links. 